### 2019-04-16 2-Chapters/5-Chapter/ALT_2018__MPBandits.git/mpBandits.tex
More on discussions to finish chapter 5

We illustrate here the empirical performances of the algorithms presented in Section~\ref{sec:5:algorithms}, used in combination with the \klUCB{} indices.
% Some plots are at pages \pageref{fig:5:MP__K9_M6_T5000_N500__4_algos__all_RegretCentralized__BayesianProblems} and \pageref{fig:5:MP__K9_M6_T10000_N1000__4_algos} and most of them in Appendix~\ref{app:5:plotsFromSec5}.

\TODOL{Include here some plots showing \RhoRand{} with \UCB{} and \klUCB{} vs \MCTopM{} with \UCB{} and \klUCB{}.}

In experiments that are not reported here, we could observe that using \klUCB{} rather than \UCB{} indices always yield better practical performance.
%, as the theoretical guarantees for single player suggested,
As the purpose of this work is not to optimize on the index policy, but rather propose new ways of using indices in a decentralized setting,
% And for some configurations, using a decentralized approach with a more efficient index family (\klUCB{} instead of \UCB) has better performances than the centralized approach.
we only report results for \klUCB.
In a first set of experiments, \MCTopM{}, \RandTopM{} and \Selfish{} are benchmarked against the state-of-the-art \RhoRand{} algorithm.
We also include a centralized multiple-play \klUCB{} algorithm,
as defined by \cite{Anantharam87a},
essentially to check that the \emph{``price of decentralized learning''} is not too large.

@@

We present results for two bandit instance: one with $K=3$ arms and means
$\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, for which two cases $M=2$ and $M=3$ are presented in Figure~\ref{fig:5:selfish_fail1}. For the second instance $K=9$ and $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$,
three cases are presented: $M=6$ in Figure~\ref{fig:5:MP__K9_M6_T10000_N1000__4_algos},
and for the two limit cases $M=2$ and $M=9=K$ in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}.
Note that here, we consider different algorithms that know the number of players $M$.
Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance.
%
We also include histograms showing the \emph{distribution} of regret at $t=T$,
as this allows to check if the regret is indeed small for \emph{each} run of the simulation.
%
For the plots showing the regret, our \emph{asymptotic} lower bound from Theorem~\ref{thm:5:BetterLowerBound} is displayed.

>>>>

We illustrate here the empirical performances of the algorithms presented in Section~\ref{sec:5:algorithms}, used in combination with the \UCB{} or \klUCB{} indices.
The analysis given in Section~\ref{sec:5:upperbounds} was focussing on \klUCB, because it is well known that it is more efficient both in practice and in theory.
To illustrate this, we first show below the result of some experiments comparing different algorithms that use either the \UCB{} or the \klUCB{} indexes.
%
Our proposed algorithms, \MCTopM{}, \RandTopM{} and \Selfish{} are benchmarked against the state-of-the-art \RhoRand{} algorithm.
We also include a centralized multiple-play \klUCB{} algorithm,
as defined by \cite{Anantharam87a},
essentially to check that the \emph{``price of decentralized learning''} is not too large.

The first experiment is considering $K=9$ arms, with means $\bm{\mu}=[0.1,0.2,\dots,0.9]$, and $M=3$ then $6$ then $9$ players.
Note that here and as in all this section we consider different algorithms that know the number of players $M$ (see Section~\ref{sub:5:unknownNumberOfPlayers} below for a discussion on the case when $M$ could be unknown).
%
Performance is measured with the \emph{expected} regret up to horizon $T=10000$, estimated based on $1000$ repetitions on the same bandit instance.
%
We report in Table~\ref{table:5:comparisonUCB_klUCB} the results in terms of mean regret, $\pm$ one standard deviation.
The conclusions are three fold: first we observe for any of the compared algorithms, using \klUCB{} is always (much) better than using \UCB.
Second we observe that our proposal \MCTopM{} is outperforming the state-of-the-art \RhoRand{} policy, and performs closely to the centralized (unrealistic) approach.
Third we observe in the last column, when $M=K=9$, that \MCTopM{} is achieving constant regret, as we proved (the regret upper-bound of Theorem~\ref{thm:5:LogarithmicRegret_MCTopMklUCB} is $\cO(1)$ if there is no arms in \Mworst{} !), and it shows that the orthogonalization scheme proposed for \MCTopM{} is very efficient, especially in comparison to previous (state-of-the-art) approaches using ranks.
%
Additional experiments and illustrations are given below.
We illustrate below the (empirical mean) regret $R_t$ as a function of time and not only results in terms of empirical mean regret $R_T$ at the end of the bandit game.

% for N in 4 100; do for M in 3 6 9; do DEBUG=True SAVEALL=False NOPLOTS=True M=$M K=30 N_JOBS=-1 N=$N T=10000 make moremultiplayers; echo "Done for N=$N and M=$M"; read; done; echo "Done for N=$N"; read; done
% TODO run the simulations with N=100 and not N=4
\begin{table}[ht]
  % \begin{footnotesize}
      \centering
      \begin{tabular}{c|ccc}
      \textbf{Algorithm} $\;$ \textbackslash $\;$ Number of players & $M=3$ & $M=6$ & $M=9$ \\
          \hline
          Centralized multiple-play \UCB{} & $317 \pm 37$ & $219 \pm 17$ & $0$ \\
          Centralized multiple-play \klUCB{} & $94 \pm 32$ & $77 \pm 7$ & $0$ \\
          \hline
          \Selfish-\UCB{} & $1175 \pm 70$ & $3705 \pm 379$ & $12709 \pm 246$ \\
          \Selfish-\klUCB{} & $230 \pm 34$ & $673 \pm 48$ & $2689 \pm 230$ \\
          \hline
          \RhoRand-\UCB{} & $1534 \pm 145$ & $4797 \pm 319$ & $12047 \pm 1056$ \\
          \RhoRand-\klUCB{} & $349 \pm 58$ & $2688 \pm 469$ & $6709 \pm 1186$ \\
          \hline
          \RandTopM-\UCB{} & $1074 \pm 36$ & $2751 \pm 229$ & $252 \pm 150$ \\
          \RandTopM-\klUCB{} & $245 \pm 30$ & $1181 \pm 100$ & $617 \pm 592$ \\
          \hline
          \MCTopM-\UCB{} & $941 \pm 23$ & $1368 \pm 86$ & $46 \pm 5$ \\
          \MCTopM-\klUCB{} & $\mathbf{277 \pm 12}$ & $\mathbf{391 \pm 70}$ & $\mathbf{50 \pm 11}$ \\
          \hline
      \end{tabular}
      \caption{Using \klUCB{} is much more efficient than using \UCB{}, for multi-player bandit (here in a simple problem with $K=9$ arms).}
      \label{table:5:comparisonUCB_klUCB}
  % \end{footnotesize}
  \end{table}

The purpose of this work is not to optimize on the index policy, but rather propose new ways of using indices in a decentralized setting,
% And for some configurations, using a decentralized approach with a more efficient index family (\klUCB{} instead of \UCB) has better performances than the centralized approach.
and as we observed that using \klUCB{} rather than \UCB{} indices always yield better practical performance,
from now on we only report results for \klUCB.
%, as the theoretical guarantees for single player suggested,

@@

We present results for two bandit instance: one with $K=3$ arms and means
$\boldsymbol{\mu} = [0.1, 0.5, 0.9]$, for which two cases $M=2$ and $M=3$ are presented in Figure~\ref{fig:5:selfish_fail1}. For the second instance $K=9$ and $\boldsymbol{\mu} = [0.1, 0.2, \dots, 0.9]$,
three cases are presented: $M=6$ in Figure~\ref{fig:5:MP__K9_M6_T10000_N1000__4_algos},
and for the two limit cases $M=2$ and $M=9=K$ in Figure~\ref{fig:5:MP__K9_M2-6-9_T10000_N200__4_algos}.
%
We also include histograms showing the \emph{distribution} of regret at $t=T$,
as this allows to check if the regret is indeed small for \emph{each} run of the simulation.
%
For the plots showing the regret, our \emph{asymptotic} lower bound from Theorem~\ref{thm:5:BetterLowerBound} is displayed.

### 2019-04-16 2-Chapters/5-Chapter/chapter5.tex
More on discussions to finish chapter 5

If the $M$ players each have a unique ID, then a simple explore-then-commit algorithm running on top of a round-Robin phase can achieve order-optimal regret.
If the players know the time horizon $T$, then fix a confidence level $\delta$. For the first $T_0$ time steps, the $M$ players will use a simple round-Robin game, using their (unique) ID to stay orthogonal: player $j$ starts at arm $j$, then $j+1$, then cycle in $\{1,\dots,K\}$. They encounter no collision in these time steps, and then user $j$ targets the $j$-th best arm among the set of $M$-best arm.
If $T_0$ is large enough, all players have built the same estimate of the ranking of the arms (and not only correctly identified the set of $M$-best arms), with high probability (at least $\delta$), and thus they will also encounter no collision and no regret from after time $T_0$.
The mean regret of such approach is easily bounded by $R_T \leq M T_0 + (1 - \delta) (T - T_0)$, so by using $\delta = 1-1/T$, $R_T = \bigO{T_0}$.
By calibrating $T_0$ based on $\delta$ (which needs prior knowledge of $T$) and the minimal gap $\Delta$ between \Mbest{} and \Mworst, one can show, using similar arguments as used by \cite{Rosenski16} for the Musical Chair policy, that with large probability a constant regret is obtained. A logarithmic regret can be obtained from the same bound, proving the order-optimality of this simple approach, if we assume that user $j$ knows it is user number $j$.
%
Without giving more details, we let the interested refer to what is is explained for the algorithms presented in \cite{DarakHanawal18,JoshiKumar2018,KumarDarak2019}.
%
These works assume a prior knowledge of $\Delta$, or other measures of the difficulty of the problem, and as such we do not.
FIXME
\TODOL{FIXME écrire ça sans être fini}

@@

The recent work \cite{BoursierPerchet18} proposed an idea they called the ``communication trick''.
It essentially allow any player to send one bit to all the others, at some pre-agreed time steps, with no modification on the model.
Thanks to this ``communication trick'', recent research efforts have been more focus on studying the basic model --without explicit communications between players--, while proposing algorithms that can rely on some communication between players.
This trick essentially use the fact that the players share a synchronized time, thus they can use a collision as a way to directly exchange one bit of information between two players.
This process is slow and not efficient, as all but two players must do nothing when player $i$ is sending a bit to player $j$, but it does build a communication protocol in the model without explicit communication between players.
This ``communication trick'' is used for instance in \cite{KaufmannAbbas19}.
We let the interested reader refer to these last two works \cite{BoursierPerchet18,KaufmannAbbas19} as they are both solid, and well explained

@@

In all the models of IoT networks considered in this thesis, we assume that \emph{the different IoT devices cannot communicate with each other}, and can only communicate with a unique gateway.
Of course, this hypothesis can be removed, and in practice in some families of wireless networks, communication between players are possible.
Note that this extension is not considering a graph of distributed agents all playing cooperatively to solve a unique bandit game, like it is studied in the ``graph bandit'' problem.
What we are interested here is an extension of our model where players can send some bits to one or all the other players, at every time steps or some time steps. Clearly, the problem is easier by allowing communication, and it is quite immediate to see that the communication capacity between players must be limited otherwise the problem becomes trivial to solve.
Indeed, imagine that at each time step $t$, all the $M$ players could share an unbounded number of bits with the other players, at no cost, even if this hypothesis is clearly unrealistic for wireless networks.
Then they can share all their observations, and they can all run the same multi-plays MAB algorithm \cite{Anantharam87a} (in a deterministic way), like for instance the extension of Thompson sampling for multi-plays studied in \cite{Komiyama15}, or extensions of \KLUCB{} from \cite{Luedtke16}.
In this setting, a logarithmic regret is obtained easily, and the regret upper-bound of the two aforementioned algorithms asymptotically achieve the lower-bound from \cite{Anantharam87a}.

An interesting extension is thus to limit the communication between players, either to a small number of bits, or just one bit.
We do not review papers that study this model, because the recent work \cite{BoursierPerchet18} introduced the ``communication trick''.
Recent research efforts have been more focus on studying the basic model --without explicit communications between players--, while proposing algorithms that can rely on some communication between players.
This ``communication trick'' is used for instance in \cite{KaufmannAbbas19}.
Due to space constraint, we let the interested reader refer to these last two works \cite{BoursierPerchet18,KaufmannAbbas19} as they are both solid, and well explained.

>>>>

If the $M$ players each have a unique ID, then a simple explore-then-commit algorithm running on top of a round-Robin phase can achieve order-optimal regret.
If the players know the time horizon $T$, then fix a confidence level $\delta$. For the first $T_0$ time steps, the $M$ players will use a simple round-Robin game, using their (unique) ID to stay orthogonal: player $j$ starts at arm $j$, then $j+1$, then cycle in $\{1,\dots,K\}$. They encounter no collision in these time steps, and then user $j$ targets the $j$-th best arm among the set of $M$-best arm.
If $T_0$ is large enough, all players have built the same estimate of the ranking of the arms (and not only correctly identified the set of $M$-best arms), with high probability (at least $\delta$), and thus they will also encounter no collision and no regret from after time $T_0$.
The mean regret of such approach is easily bounded by $R_T \leq M T_0 + (1 - \delta) (T - T_0)$, so by using $\delta = 1-1/T$, $R_T = \bigO{T_0}$.
By calibrating $T_0$ based on $\delta$ (which needs prior knowledge of $T$) and the minimal gap $\Delta$ between \Mbest{} and \Mworst, one can show, using similar arguments as used by \cite{Rosenski16} for the Musical Chair policy, that with large probability a constant regret is obtained. A logarithmic regret can be obtained from the same bound, proving the order-optimality of this simple approach, if we assume that user $j$ knows it is user number $j$.
%
Without giving more details, we let the interested refer to what is is explained for the algorithms presented in \cite{DarakHanawal18,JoshiKumar2018,KumarDarak2019}.
%
These works assume a prior knowledge of $\Delta$, or other measures of the difficulty of the problem, and as such we do not.
FIXME
\TODOL{FIXME écrire ça sans être méchant envers leurs travaux !}

@@

The recent work \cite{BoursierPerchet18} proposed an idea they called the ``communication trick''.
It essentially allow any player to send one bit to all the others, at some pre-agreed time steps, with no modification on the model.
Thanks to this ``communication trick'', recent research efforts have been more focus on studying the basic model --without explicit communications between players--, while proposing algorithms that can rely on some communication between players.
This trick essentially use the fact that the players share a synchronized time, thus they can use a collision as a way to directly exchange one bit of information between two players.
This process is slow and not efficient, as all but two players must do nothing when player $i$ is sending a bit to player $j$, but it does build a communication protocol in the model without explicit communication between players.
This ``communication trick'' is used for instance in \cite{KaufmannAbbas19}.
We let the interested reader refer to these last two works \cite{BoursierPerchet18,KaufmannAbbas19} as they are both solid, and well explained.

@@

In all the models of IoT networks considered in this thesis, we assume that \emph{the different IoT devices cannot communicate with each other}, and can only communicate with a unique gateway.
Of course, this hypothesis can be removed, and in practice in some families of wireless networks, communication between players are possible.
Note that this extension is not considering a graph of distributed agents all playing cooperatively to solve a unique bandit game, like it is studied in the ``graph bandit'' problem.
What we are interested here is an extension of our model where players can send some bits to one or all the other players, at every time steps or some time steps.

Clearly, the problem is easier by allowing communication, and it is quite immediate to see that the communication capacity between players must be limited otherwise the problem is already known and solved.
Indeed, imagine that at each time step $t$, all the $M$ players could share an unbounded number of bits with the other players, at no cost, even if this hypothesis is clearly unrealistic for wireless networks.
Then they can share all their observations, and they can all run the same multiple-plays MAB algorithm \cite{Anantharam87a} (in a deterministic way), like for instance the extension of Thompson sampling for multiple-plays studied in \cite{Komiyama15}, or extensions of \KLUCB{} from \cite{Luedtke16}.
In this setting, a logarithmic regret is obtained easily, and the regret upper-bound of the two aforementioned algorithms asymptotically achieve the lower-bound from \cite{Anantharam87a}.

An interesting extension is thus to limit the communication between players, either to a small number of bits, or just one bit, at every or only some time steps.
We identified that the state-of-the-art on this direction of research consists in the two very recent works published in April 2019, while I was finishing this chapter: \cite{tao2019collaborative} and \cite{wang2019distributed}.
Distributed pure exploration is studied in \cite{tao2019collaborative}, where the proposed new lower bounds for the regret of any algorithm for the distributed best arm identification problem, under the fixed time or fixed confidence settings.
The also propose effective algorithms that asymptotically match their lower-bounds (up-to logarithmic factors, in some cases).
The second work studies distributed learning for (not necessarily stochastic) multi-armed bandits as well as linear bandits \cite{tao2019collaborative}.
For a distributed $K$-armed bandit with $M$ agents, they developed two protocols achieving near-optimal regret $\cO(\sqrt{M K T \log(T)})$,
and requiring little communication cost,
one is independent of the time horizon $T$ and use $\cO(M \log(T))$ communications,
and the other is independent of the number of arms $K$ and use $\cO(M K \log(M))$ communications.
Their model and algorithms fit in the distributed, decentralized framework we advertise in all this thesis, and this last work is very interesting.

Additionally, after the recent work \cite{BoursierPerchet18} introduced the ``communication trick'',
some recent research efforts have been more focus on studying the basic model --without explicit communications between players--, while proposing algorithms that can rely on some communication between players.
This ``communication trick'' is used for instance in \cite{KaufmannAbbas19}.
Due to space constraint, we let the interested reader refer to these last two works \cite{BoursierPerchet18,KaufmannAbbas19} as they are both solid, and well explained.
